WDIAS : A Microservices-Based Weather Data Integration and Assimilation System\\

### 2020-05-30 method/sec_6-query.tex
4.2 Done
Work load done

We used the adapter-query primarily for supporting geo-spatial queries. As the primary service for searching for timeseries, it supports retrieval of timeseries metadata for external queries as well. The Appendix B described the search query capabilities of \acrshort{wdias}.
The system provides timeseries metadata query endpoints to search for existing timeseries metadata within the system using primary key attributes.
Also, it provides a set of geo-query endpoints, and users can perform geo-search queries with sending geo-information in geoJson format. The following geo-queries are supported by the system, and a detailed description appears in Appendix B.
\begin{itemize}
    \item Query Locations within Area
    \item Query parameters in Locations
    \item Query Timeseries by Location
    \item Query Timeseries by Locations
    \item Query Timeseries by Parameter
    \item Query Timeseries in Area
    \item Query Timeseries in Area by Parameter
    \item Query All Timeseries
\end{itemize}

>>>>

We used the query adapter primarily for supporting geo-spatial queries. As the primary service for searching for timeseries, it supports retrieval of timeseries metadata for external queries as well. The Appendix B described the search query capabilities within \acrshort{wdias}.

The system provides timeseries metadata query endpoints to search for existing timeseries metadata within the system using primary key attributes.
Also, it provides a set of geo-query endpoints, and users can perform geo-search queries with sending geo-information in geoJson format. The following geo-queries are supported by the system, and a detailed description appears in Appendix B.
\begin{itemize}
    \item Query locations within area
    \item Query parameters in locations
    \item Query timeseries by location
    \item Query timeseries by locations
    \item Query timeseries by Parameter
    \item Query timeseries in area
    \item Query timeseries in area by parameter
    \item Query all timeseries
\end{itemize}

### 2020-05-30 results/sec_2-workload.tex
Work load done

To create a higher workload, we used JMeter's Distributed Testing capabilities. JMeter uses the master-slave approach for distributed testing and allows us to handle all the test cases via a single master instance. However, it has the limitation of \emph{the same test plan runs by all the servers} rather than distributing the work among slave instances. In other words, JMeter does not distribute the load between servers, and each slave service runs the full test plan. So if we set 1,000 threads per test plan and run the test using six JMeter servers, the distributed test setup ends up injecting 6,000 threads. During JMeter distributed mode, the master server triggers the same copy of the test plan on configured slave nodes in parallel. After successfully running the test plans, the master server gathers results from slave servers.

@@

While implementing JMeter test plans, it uses threads to simulate the users. The root component of a test plan is the thread group. There are different kinds of thread groups available such as thread group (basic thread group), arrivals thread Group, free form arrivals thread group, stepping thread group, concurrency thread group, and ultimate thread group.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Closed vs. open workload models}
\label{subse:closed_vs_open_workload}

\begin{itemize}
    \item \emph{closed system model} \cite{Haggett1998AnWales} -- a new request is only triggered by the completion of a previous request, following by a think time. The system has negative feedback that makes it impossible to bury-out the service, so users wait for the responses before making new requests.
    \item \emph{open system model} -- new requests arrival independently of completions, e.g., according to a stochastic process or fixed trace. The system has no negative feedback.
\end{itemize}

While implementing the test cases for \acrshort{wdias}, we used the concurrency thread group with the throughput shaping timer extension of JMeter because it supports the open workload approach. Using other thread groups, we are required to find the exact number of threads and timer delays that produce the desired number of \acrfull{rps} to the server, which is known as the \emph{closed workload}. By using the \emph{throughput shaping timer}, we only need to configure the \acrshort{rps} throughout the test plan. Then the throughput shaping timer provides a feedback loop to the concurrency thread group via schedule feedback function to dynamically maintain thread count required to achieve target \acrshort{rps} \cite{KarunarathneGihanWdias/wdias-performance-test:JMeter.}.

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{results/work_load/test_prod_throughtput_shaping_timer.png}
    \caption{\acrshort{wdias} load testing throughput shaping timer \acrshort{rps} configurations}
    \label{fi:test_prod_throughtput_shaping_timer}
\end{figure}

@@

\begin{itemize}
    \item HTTP Request Defaults
    \item HTTP Header Manager
    \item Timeseries Metadata CSV
    \item If Scalar or Vector
        \begin{itemize}
            \item Insert Timeseries
        \end{itemize}
    \item If a Grid
    \begin{itemize}
        \item Insert Grid Timeseries
    \end{itemize}
    \item Wait for extensions timeseries
    \item Interest Area preprocessor
    \item Query If Scalar Or Vector
    \item Retrieve If Scalar or Vector
        \begin{itemize}
            \item Retrieve Timeseries
        \end{itemize}
    \item Retrieve If a Grid
    \begin{itemize}
        \item Retrieve Grid Timeseries
    \end{itemize}
    \item Increment date
    \item Throughput shaping timer prod
\end{itemize}

@@

\begin{itemize}
    \item HTTP Request Defaults
    \item HTTP Header Manager
    \item Locations CSV
    \item Interest Area preprocessor
    \begin{itemize}
        \item /Location: * → Locations
    	\item /Location: Area → Locations
    	\item /Parameter: Location → Parameters
    	\item /Parameter: Locations → Parameters
    	\item /Timeseries: Location → Timeseries
    	\item /Timeseries: Locations → Timeseries
    	\item /Timeseries: Locations, Parameter → Timeseries
    	\item /Timeseries: Area → Timeseries
    	\item /Timeseries: Area, Parameter → Timeseries
    	\item /Timeseries: *, Parameter → Timeseries
    	\item /Timeseries: * → Timeseries
	\end{itemize}
\end{itemize}

@@

\subsubsection{JMeter performance tuning}
\begin{itemize}
    \item Use CSV files rather than random number generation
    \item Use JSR223 scripts with Groovy for data processing for test cases
    \item Using Throughput Shaping Timer to get more accurate \acrshort{rps}
\end{itemize}

@@

The \acrshort{wdias} uses Helm \cite{CNCFHelmDocs} which is a package manager for \acrshort{k8s}. Each Also, we used Helm \cite{CNCFHelmDocs} that is a package manager for \acrshort{k8s}. Each microservice within the system deploys and maintains as a Helm chart and the Helm charts \cite{KarunarathneWdias-helm-charts:Deployments} for each microservice available within the codebase. For the databases, we used the official helm charts. Using Helm, we can configure resources, the number of replicas needed, deploy new changes, and manage credentials, storage, and proxy settings that are needed for each microservice.

>>>>

To create a higher workload, we used JMeter's distributed testing capabilities. JMeter uses the master-slave approach for distributed testing and allows us to handle all the test cases via a single master instance. However, it has the limitation of \emph{the same test plan runs by all the servers} rather than distributing the work among slave instances. In other words, JMeter does not distribute the load between servers, and each slave service runs the full test plan. So, if we set 1,000 threads per test plan and run the test using six JMeter servers, the distributed test setup ends up injecting 6,000 threads. During JMeter distributed mode, the master server triggers the same copy of the test plan on configured slave nodes in parallel. After successfully running the test plans, the master server gathers results from slave servers.

@@

While implementing JMeter test plans, it uses threads to simulate the users. The root component of a test plan is the thread group. There are different kinds of thread groups available such as thread group (basic thread group), arrivals thread group, free form arrivals thread group, stepping thread group, concurrency thread group, and ultimate thread group.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Closed vs. Open Workload Models}
\label{subse:closed_vs_open_workload}

\begin{itemize}
    \item \emph{closed system model} \cite{Haggett1998AnWales} -- A new request is only triggered by the completion of a previous request, following by a think time. The system has negative feedback that makes it impossible to bury-out the service, so users wait for the responses before making new requests.
    \item \emph{open system model} -- New requests arrival independently of completions, e.g., according to a stochastic process or fixed trace. The system has no negative feedback.
\end{itemize}

While implementing the test cases for \acrshort{wdias}, we used the concurrency thread group with the throughput shaping timer extension of JMeter because it supports the open workload approach. Using other thread groups, we are required to find the exact number of threads and timer delays that produce the desired number of \acrshort{rps} to the server, which is known as the \emph{closed workload}. By using the \emph{throughput shaping timer}, we only need to configure the \acrshort{rps} throughout the test plan. Then the throughput shaping timer provides a feedback loop to the concurrency thread group via schedule feedback function to dynamically maintain thread count required to achieve target \acrshort{rps} \cite{KarunarathneGihanWdias/wdias-performance-test:JMeter.}.

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{results/work_load/test_prod_throughtput_shaping_timer.png}
    \caption{\acrshort{wdias} load testing throughput shaping timer \acrshort{rps} configurations.}
    \label{fi:test_prod_throughtput_shaping_timer}
\end{figure}

@@

\begin{itemize}
    \item HTTP request defaults
    \item HTTP header manager
    \item Timeseries metadata CSV
    \item If scalar or vector
        \begin{itemize}
            \item Insert timeseries
        \end{itemize}
    \item If a grid
    \begin{itemize}
        \item Insert grid timeseries
    \end{itemize}
    \item Wait for extensions timeseries
    \item Interest area preprocessor
    \item Query if scalar or vector
    \item Retrieve if scalar or vector
        \begin{itemize}
            \item Retrieve timeseries
        \end{itemize}
    \item Retrieve if a grid
    \begin{itemize}
        \item Retrieve grid timeseries
    \end{itemize}
    \item Increment date
    \item Throughput shaping timer prod
\end{itemize}

@@

\begin{itemize}
    \item HTTP request defaults
    \item HTTP header manager
    \item Locations CSV
    \item Interest area preprocessor
    \begin{itemize}
        \item /Location: * → Locations
    	\item /Location: Area → Locations
    	\item /Parameter: Location → Parameters
    	\item /Parameter: Locations → Parameters
    	\item /Timeseries: Location → Timeseries
    	\item /Timeseries: Locations → Timeseries
    	\item /Timeseries: Locations, Parameter → Timeseries
    	\item /Timeseries: Area → Timeseries
    	\item /Timeseries: Area, Parameter → Timeseries
    	\item /Timeseries: *, Parameter → Timeseries
    	\item /Timeseries: * → Timeseries
	\end{itemize}
\end{itemize}

@@

\subsubsection{JMeter performance tuning}
\begin{itemize}
    \item Use CSV files rather than random number generation
    \item Use JSR223 scripts with Groovy for data processing for test cases
    \item Using throughput shaping timer to get more accurate \acrshort{rps}
\end{itemize}

@@

The \acrshort{wdias} uses Helm \cite{CNCFHelmDocs} which is a package manager for \acrshort{k8s}. Each microservice within the system deploys and maintains as a Helm chart and the Helm charts \cite{KarunarathneWdias-helm-charts:Deployments} for each microservice available within the codebase. For the databases, we used the official helm charts. Using Helm, we can configure resources, the number of replicas needed, deploy new changes, and manage credentials, storage, and proxy settings that are needed for each microservice.

### 2020-05-30 results/sec_3-observations.tex
Work load done

This section includes the observations based on the test plans discussed in Section \cref{se:test_plan}, and the experimental system setup explained in \cref{se:workload}. Sections \ref{subse:obs_test_plan_all_60min} to \ref{subse:obs_test_plan_all_auto_15min}, we discuss the observations collected with performing the Load test plans over 60 minutes data (24 data points per request), 30 minutes data (48 data points per request), and 15 minutes data (96 data points per request) by varying the request size. Then we perform another round of 15 minutes data with enabling auto-scaling for the higher resource utilized microservices.

@@

\begin{table}[ht]
\caption{Throughput and Latency of load testing with 60-min data}
\footnotesize
\begin{tabulary}{\linewidth}{|L|R|R|R|R|R|R|R|R|}
\hline
\textbf{Label} & \textbf{Samples} & \textbf{Avg} & \textbf{Min} & \textbf{Max} & \textbf{90\% Line} & \textbf{Std.Dev.} & \textbf{Error} & \textbf{RPS} \\ \hline
Insert Timeseries & 71826 & 28 & 13 & 2773 & 31 & 58.74 & 0.00\% & 40.5 \\ \hline
Retrieve Timeseries & 71796 & 8 & 7 & 242 & 10 & 4.18 & 0.00\% & 40.7 \\ \hline
Insert Grid & 7982 & 23 & 21 & 126 & 26 & 4.23 & 0.06\% & 4.5 \\ \hline
Retrieve Grid & 7979 & 68 & 59 & 238 & 75 & 10.11 & 0.00\% & 4.5 \\ \hline
Query: Location & 71804 & 3 & 2 & 109 & 3 & 1.52 & 0.00\% & 40.5 \\ \hline
\textbf{TOTAL} & 311182 & 127 & 0 & 2773 & 503 & 207.80 & 0.00\% & 175.4 \\ \hline
\end{tabulary}
\label{tab:obs_all_60_min_summary}
\end{table}
Insert timeseries performed over scalar and vector data types with each request sent with 24 data points. On average, it took 28 milliseconds to insert scalar and vector data into the \acrshort{wdias}, and up to 90\% percentile inserted data within 31ms. Noticeably it has 0\% of errors, which means that all the requests completed successfully. Also, \acrshort{wdias} handled 40.5 \acrshort{rps} of scalar and vector data requests with the given workload. The retrieval of scalar and vector timeseries has lesser delay than the insertion.
On the other hand, insert grid timeseries data performed within 23ms, and almost all the requests perform within the same amount of time. Hence, the standard deviation is smaller. Delay with inserting grid timeseries smaller because those requests are handled asynchronously. Retrieve grid timeseries data performed with a bit higher value of latency of 68 milliseconds due to performed the retrieve request on-demand. Both insert and retrieve grid timeseries data have 4.5 \acrshort{rps} since it only gets 10\% of the number of requests per given time during the load test.

@@

Next, we performed the test plan with 30-minutes resolution data, which means 48 data points per each request for the scalar and vector data types, and 48 ASCII grid files per each insertion request for the grid data type. Also, the test plan performed round up to 311k number of sample requests, which is almost similar to the number of samples processed with hourly resolution data.

@@

\cref{tab:obs_all_30_min_summary}shows the response latency summary details and \acrshort{rps} with 30-min resolution data. The results are almost similar to the observations in \cref{subse:obs_test_plan_all_60min}. Even though we doubled the request size, \acrshort{wdias} was able to keep the performance almost the same instead of increasing the latency twice.

@@

\cref{fi:test_obs_all_30m_response_vs_threads} shows the latency against the number of active threads for the test plan with 30-min resolution data. As the graph shows, the response time kept almost constant while increasing the number of active threads. Thus, we can say that \acrshort{wdias} is a scalable system since it was able to increase the throughput without a significant change in the latency. When the number of active threads increased, \cref{fi:test_obs_all_30m_response_vs_threads} shows uncertainty in the insertion and retrieval of scalar and vector data. Some of the facts cause this is, those operations are handled on-demand. At the same time, the system is writing timeseries while reading from the timeseries database. The InfluxDB can fine-tune for writes or read based on the requirements, but we used it with default configurations.

>>>>

This section includes the observations based on the test plans discussed in \cref{se:test_plan}, and the experimental system setup explained in \cref{se:workload}. Sections \ref{subse:obs_test_plan_all_60min} to \ref{subse:obs_test_plan_all_auto_15min}, we discuss the observations collected with performing the load test plans over 60 minutes data (24 data points per request), 30 minutes data (48 data points per request), and 15 minutes data (96 data points per request) by varying the request size. Then we perform another round of 15 minutes data with enabling auto-scaling for the higher resource utilized microservices.

@@

\begin{table}[ht]
\caption{Throughput and Latency of load testing with 60-min data}
\footnotesize
\begin{tabulary}{\linewidth}{|L|R|R|R|R|R|R|R|R|}
\hline
\textbf{Label} & \textbf{Samples} & \textbf{Avg} & \textbf{Min} & \textbf{Max} & \textbf{90\% Line} & \textbf{Std.Dev.} & \textbf{Error} & \textbf{RPS} \\ \hline
Insert Timeseries & 71826 & 28 & 13 & 2773 & 31 & 58.74 & 0.00\% & 40.5 \\ \hline
Retrieve Timeseries & 71796 & 8 & 7 & 242 & 10 & 4.18 & 0.00\% & 40.7 \\ \hline
Insert Grid & 7982 & 23 & 21 & 126 & 26 & 4.23 & 0.06\% & 4.5 \\ \hline
Retrieve Grid & 7979 & 68 & 59 & 238 & 75 & 10.11 & 0.00\% & 4.5 \\ \hline
Query: Location & 71804 & 3 & 2 & 109 & 3 & 1.52 & 0.00\% & 40.5 \\ \hline
\textbf{TOTAL} & 311182 & 127 & 0 & 2773 & 503 & 207.80 & 0.00\% & 175.4 \\ \hline
\end{tabulary}
\label{tab:obs_all_60_min_summary}
\end{table}
Insert timeseries performed over scalar and vector data types with each request sent with 24 data points. On average, it took 28 milliseconds to insert scalar and vector data into the \acrshort{wdias}, and up to 90\% percentile inserted data within 31 milliseconds. Noticeably it has 0\% of errors, which means that all the requests completed successfully. Also, \acrshort{wdias} handled 40.5 \acrshort{rps} of scalar and vector data requests with the given workload. The retrieval of scalar and vector timeseries has lesser delay than the insertion.
On the other hand, insert grid timeseries data performed within 23 milliseconds, and almost all the requests perform within the same amount of time. Hence, the standard deviation is smaller. Delay with inserting grid timeseries smaller because those requests are handled asynchronously. Retrieve grid timeseries data performed with a bit higher value of latency of 68 milliseconds due to performed the retrieve request on-demand. Both insert and retrieve grid timeseries data have 4.5 \acrshort{rps} since it only gets 10\% of the number of requests per given time during the load test.

@@

Next, we performed the test plan with 30-minute resolution data, which means 48 data points per each request for the scalar and vector data types, and 48 ASCII grid files per each insertion request for the grid data type. Also, the test plan performed round up to 311k number of sample requests, which is almost similar to the number of samples processed with hourly resolution data.

@@

\cref{tab:obs_all_30_min_summary} shows the response latency summary details and \acrshort{rps} with 30-minute resolution data. The results are almost similar to the observations in \cref{subse:obs_test_plan_all_60min}. Even though we doubled the request size, \acrshort{wdias} was able to keep the performance almost the same instead of increasing the latency twice.

@@

\cref{fi:test_obs_all_30m_response_vs_threads} shows the latency against the number of active threads for the test plan with 30-minute resolution data. As the graph shows, the response time kept almost constant while increasing the number of active threads. Thus, we can say that \acrshort{wdias} is a scalable system since it was able to increase the throughput without a significant change in the latency. When the number of active threads increased, \cref{fi:test_obs_all_30m_response_vs_threads} shows uncertainty in the insertion and retrieval of scalar and vector data. Some of the facts cause this is, those operations are handled on-demand. At the same time, the system is writing timeseries while reading from the timeseries database. The InfluxDB can fine-tune for writes or read based on the requirements, but we used it with default configurations.
